---
title: "LDA"
output: html_document
---

```{r, message=FALSE, warning=FALSE}
packages.used=c("rvest", "tibble", "qdap", 
                "sentimentr", "gplots", "dplyr",
                "tm", "syuzhet", "factoextra", 
                "beeswarm", "scales", "RColorBrewer",
                "RANN", "tm", "topicmodels")

# check packages that need to be installed.
packages.needed=setdiff(packages.used, 
                        intersect(installed.packages()[,1], 
                                  packages.used))
# install additional packages
if(length(packages.needed)>0){
  install.packages(packages.needed, dependencies = TRUE)
}

# load packages
library("rvest")
library("tibble")
library("qdap")
library("sentimentr")
library("gplots")
library("dplyr")
library("tm")
library("syuzhet")
library("factoextra")
library("beeswarm")
library("scales")
library("RColorBrewer")
library("RANN")
library("tm")
library("topicmodels")

source("../lib/plotstacked.R")
source("../lib/speechFuncs.R")
```

#Data harvest: scrap speech URLs from <http://www.presidency.ucsb.edu/>.
```{r, message=FALSE, warning=FALSE}
### Inauguaral speeches
main.page <- read_html(x = "http://www.presidency.ucsb.edu/inaugurals.php")
# Get link URLs
# f.speechlinks is a function for extracting links from the list of speeches. 
inaug=f.speechlinks(main.page)
#head(inaug)
as.Date(inaug[,1], format="%B %e, %Y")
inaug=inaug[-nrow(inaug),] # remove the last line, irrelevant due to error.

#### Nomination speeches
main.page=read_html("http://www.presidency.ucsb.edu/nomination.php")
# Get link URLs
nomin <- f.speechlinks(main.page)
#head(nomin)
#
#### Farewell speeches
main.page=read_html("http://www.presidency.ucsb.edu/farewell_addresses.php")
# Get link URLs
farewell <- f.speechlinks(main.page)
#head(farewell)
```

#Using speech metadata posted on <http://www.presidency.ucsb.edu/>,prepared CSV data sets for the speeches we will scrap. 

```{r}
inaug.list=read.csv("../data/inauglist.csv", stringsAsFactors = FALSE)
nomin.list=read.csv("../data/nominlist.csv", stringsAsFactors = FALSE)
farewell.list=read.csv("../data/farewelllist.csv", stringsAsFactors = FALSE)
```

#scrap the texts of speeches from the speech URLs.

```{r}
speech.list=rbind(inaug.list, nomin.list, farewell.list)
speech.list$type=c(rep("inaug", nrow(inaug.list)),
                   rep("nomin", nrow(nomin.list)),
                   rep("farewell", nrow(farewell.list)))
speech.url=rbind(inaug, nomin, farewell)
speech.list=cbind(speech.list, speech.url)
```



```{r}
# Loop over each row in speech.list
speech.list$fulltext=NA
for(i in seq(nrow(speech.list))) {
  text <- read_html(speech.list$urls[i]) %>% # load the page
    html_nodes(".displaytext") %>% # isloate the text
    html_text() # get the text
  speech.list$fulltext[i]=text
  # Create the file name
  filename <- paste0("../data/fulltext/", 
                     speech.list$type[i],
                     speech.list$File[i], "-", 
                     speech.list$Term[i], ".txt")
  sink(file = filename) %>% # open file to write 
  cat(text)  # write the file
  sink() # close the file
}
```

Trump, as president-elect that has not been a politician, do not have a lot of formal speeches yet. For our textual analysis, we manually add several public transcripts from Trump:
+ [Transcript: Donald Trump's full immigration speech, annotated. LA Times, 08/31/2016] (http://www.latimes.com/politics/la-na-pol-donald-trump-immigration-speech-transcript-20160831-snap-htmlstory.html)
+ [Transcript of Donald Trump’s speech on national security in Philadelphia
- The Hill, 09/07/16](http://thehill.com/blogs/pundits-blog/campaign/294817-transcript-of-donald-trumps-speech-on-national-security-in)
+ [Transcript of President-elect Trump's news conference
CNBC, 01/11/2017](http://www.cnbc.com/2017/01/11/transcript-of-president-elect-donald-j-trumps-news-conference.html)

```{r}
speech1=paste(readLines("../data/fulltext/SpeechDonaldTrump-NA.txt", 
                  n=-1, skipNul=TRUE),
              collapse=" ")
speech2=paste(readLines("../data/fulltext/SpeechDonaldTrump-NA2.txt", 
                  n=-1, skipNul=TRUE),
              collapse=" ")
speech3=paste(readLines("../data/fulltext/PressDonaldTrump-NA.txt", 
                  n=-1, skipNul=TRUE),
              collapse=" ")

Trump.speeches=data.frame(
  President=rep("Donald J. Trump", 3),
  File=rep("DonaldJTrump", 3),
  Term=rep(0, 3),
  Party=rep("Republican", 3),
  Date=c("August 31, 2016", "September 7, 2016", "January 11, 2017"),
  Words=c(word_count(speech1), word_count(speech2), word_count(speech3)),
  Win=rep("yes", 3),
  type=rep("speeches", 3),
  links=rep(NA, 3),
  urls=rep(NA, 3),
  fulltext=c(speech1, speech2, speech3)
)

speech.list=rbind(speech.list, Trump.speeches)
```

#data Processing 
```{r, message=FALSE, warning=FALSE}
sentence.list=NULL
for(i in 1:nrow(speech.list)){
  sentences=sent_detect(speech.list$fulltext[i],
                        endmarks = c("?", ".", "!", "|",";"))
  if(length(sentences)>0){
    emotions=get_nrc_sentiment(sentences)
    word.count=word_count(sentences)
    # colnames(emotions)=paste0("emo.", colnames(emotions))
    # in case the word counts are zeros?
    emotions=diag(1/(word.count+0.01))%*%as.matrix(emotions)
    sentence.list=rbind(sentence.list, 
                        cbind(speech.list[i,-ncol(speech.list)],
                              sentences=as.character(sentences), 
                              word.count,
                              emotions,
                              sent.id=1:length(sentences)
                              )
    )
  }
}
```

Some non-sentences exist in raw data due to erroneous extra end-of sentence marks. 
```{r}
sentence.list=
  sentence.list%>%
  filter(!is.na(word.count)) 

```


#Topic modeling

For topic modeling, we prepare a corpus of sentence snipets as follows. For each speech, we start with sentences and prepare a snipet with a given sentence with the flanking sentences. 

```{r}
corpus.list=sentence.list[2:(nrow(sentence.list)-1), ]
sentence.pre=sentence.list$sentences[1:(nrow(sentence.list)-2)]
sentence.post=sentence.list$sentences[3:(nrow(sentence.list)-1)]
corpus.list$snipets=paste(sentence.pre, corpus.list$sentences, sentence.post, sep=" ")
rm.rows=(1:nrow(corpus.list))[corpus.list$sent.id==1]
rm.rows=c(rm.rows, rm.rows-1)
corpus.list=corpus.list[-rm.rows, ]
```

## Text mining
```{r}
docs <- Corpus(VectorSource(corpus.list$snipets))
writeLines(as.character(docs[[sample(1:nrow(corpus.list), 1)]]))
```

### Text basic processing
Adapted from <https://eight2late.wordpress.com/2015/09/29/a-gentle-introduction-to-topic-modeling-using-r/>.

```{r}
#remove potentially problematic symbols
docs <-tm_map(docs,content_transformer(tolower))
writeLines(as.character(docs[[sample(1:nrow(corpus.list), 1)]]))

#remove punctuation
docs <- tm_map(docs, removePunctuation)
writeLines(as.character(docs[[sample(1:nrow(corpus.list), 1)]]))

#Strip digits
docs <- tm_map(docs, removeNumbers)
writeLines(as.character(docs[[sample(1:nrow(corpus.list), 1)]]))

#remove stopwords
docs <- tm_map(docs, removeWords, stopwords("english"))
writeLines(as.character(docs[[sample(1:nrow(corpus.list), 1)]]))

#remove whitespace
docs <- tm_map(docs, stripWhitespace)
writeLines(as.character(docs[[sample(1:nrow(corpus.list), 1)]]))

#Stem document
docs <- tm_map(docs,stemDocument)
writeLines(as.character(docs[[sample(1:nrow(corpus.list), 1)]]))
```
Gengerate document-term matrices. 

```{r}
dtm <- DocumentTermMatrix(docs)
#convert rownames to filenames#convert rownames to filenames
rownames(dtm) <- paste(corpus.list$type, corpus.list$File,
                       corpus.list$Term, corpus.list$sent.id, sep="_")

rowTotals <- apply(dtm , 1, sum) #Find the sum of words in each Document

dtm  <- dtm[rowTotals> 0, ]
corpus.list=corpus.list[rowTotals>0, ]

```

Run LDA

```{r}
#Set parameters for Gibbs sampling
burnin <- 4000
iter <- 2000
thin <- 500
seed <-list(2003,5,63,100001,765)
nstart <- 5
best <- TRUE

#Number of topics
k <- 15

#Run LDA using Gibbs sampling
ldaOut <-LDA(dtm, k, method="Gibbs", control=list(nstart=nstart, 
                                                 seed = seed, best=best,
                                                 burnin = burnin, iter = iter, 
                                                 thin=thin))
#write out results
#docs to topics
ldaOut.topics <- as.matrix(topics(ldaOut))
table(c(1:k, ldaOut.topics))
write.csv(ldaOut.topics,file=paste("../out/LDAGibbs",k,"DocsToTopics.csv"))

#top 6 terms in each topic
ldaOut.terms <- as.matrix(terms(ldaOut,20))
write.csv(ldaOut.terms,file=paste("../out/LDAGibbs",k,"TopicsToTerms.csv"))

#probabilities associated with each topic assignment
topicProbabilities <- as.data.frame(ldaOut@gamma)
write.csv(topicProbabilities,file=paste("../out/LDAGibbs",k,"TopicProbabilities.csv"))
```
```{r}
terms.beta=ldaOut@beta
terms.beta=scale(terms.beta)
topics.terms=NULL
for(i in 1:k){
  topics.terms=rbind(topics.terms, ldaOut@terms[order(terms.beta[i,], decreasing = TRUE)[1:7]])
}
topics.terms
ldaOut.terms
```

Based on the most popular terms and the most salient terms for each topic, we assign a hashtag to each topic. This part require manual setup as the topics are likely to change. 

```{r}
topics.hash=c("Economy", "America", "Defense", "Belief", "Election", "Patriotism", "Unity", "Government", "Reform", "Temporal", "WorkingFamilies", "Freedom", "Equality", "Misc", "Legislation")
corpus.list$ldatopic=as.vector(ldaOut.topics)
corpus.list$ldahash=topics.hash[ldaOut.topics]

colnames(topicProbabilities)=topics.hash
corpus.list.df=cbind(corpus.list, topicProbabilities)
```

## Clustering of topics
```{r, fig.width=3, fig.height=4}
par(mar=c(1,1,1,1))
topic.summary=tbl_df(corpus.list.df)%>%
              filter(type%in%c("nomin", "inaug"), File%in%sel.comparison)%>%
              select(File, Economy:Legislation)%>%
              group_by(File)%>%
              summarise_each(funs(mean))
topic.summary=as.data.frame(topic.summary)
rownames(topic.summary)=topic.summary[,1]

# [1] "Economy"         "America"         "Defense"         "Belief"         
# [5] "Election"        "Patriotism"      "Unity"           "Government"     
# [9] "Reform"          "Temporal"        "WorkingFamilies" "Freedom"        
# [13] "Equality"        "Misc"            "Legislation"       

topic.plot=c(1, 13, 9, 11, 8, 3, 7)
print(topics.hash[topic.plot])

heatmap.2(as.matrix(topic.summary[,topic.plot+1]), 
          scale = "column", key=F, 
          col = bluered(100),
          cexRow = 0.9, cexCol = 0.9, margins = c(8, 8),
          trace = "none", density.info = "none")
```

```{r, fig.width=3.3, fig.height=5}
# [1] "Economy"         "America"         "Defense"         "Belief"         
# [5] "Election"        "Patriotism"      "Unity"           "Government"     
# [9] "Reform"          "Temporal"        "WorkingFamilies" "Freedom"        
# [13] "Equality"        "Misc"            "Legislation"       
 

par(mfrow=c(5, 1), mar=c(1,1,2,0), bty="n", xaxt="n", yaxt="n")

topic.plot=c(1, 13, 14, 15, 8, 9, 12)
print(topics.hash[topic.plot])

speech.df=tbl_df(corpus.list.df)%>%filter(File=="GeorgeBush", type=="nomin",Term==1)%>%select(sent.id, Economy:Legislation)
speech.df=as.matrix(speech.df)
speech.df[,-1]=replace(speech.df[,-1], speech.df[,-1]<1/15, 0.001)
speech.df[,-1]=f.smooth.topic(x=speech.df[,1], y=speech.df[,-1])
plot.stacked(speech.df[,1], speech.df[,topic.plot+1], 
             xlab="Sentences", ylab="Topic share", main="George Bush, Nomination")

speech.df=tbl_df(corpus.list.df)%>%filter(File=="WilliamJClinton", type=="nomin", Term==1)%>%select(sent.id, Economy:Legislation)
speech.df=as.matrix(speech.df)
speech.df[,-1]=replace(speech.df[,-1], speech.df[,-1]<1/15, 0.001)
speech.df[,-1]=f.smooth.topic(x=speech.df[,1], y=speech.df[,-1])
plot.stacked(speech.df[,1], speech.df[,topic.plot+1],
            xlab="Sentences", ylab="Topic share", main="Bill Clinton, Nomination")

speech.df=tbl_df(corpus.list.df)%>%filter(File=="GeorgeWBush", type=="nomin", Term==1)%>%select(sent.id, Economy:Legislation)
speech.df=as.matrix(speech.df)
speech.df[,-1]=replace(speech.df[,-1], speech.df[,-1]<1/15, 0.001)
speech.df[,-1]=f.smooth.topic(x=speech.df[,1], y=speech.df[,-1])
plot.stacked(speech.df[,1], speech.df[,topic.plot+1], 
            xlab="Sentences", ylab="Topic share", main="George W Bush, Nomination")

speech.df=tbl_df(corpus.list.df)%>%filter(File=="BarackObama", type=="nomin", Term==1)%>%select(sent.id, Economy:Legislation)
speech.df=as.matrix(speech.df)
speech.df[,-1]=replace(speech.df[,-1], speech.df[,-1]<1/15, 0.001)
speech.df[,-1]=f.smooth.topic(x=speech.df[,1], y=speech.df[,-1])
plot.stacked(speech.df[,1], speech.df[,topic.plot+1],
            xlab="Sentences", ylab="Topic share", main="Barack Obama, Nomination")

speech.df=tbl_df(corpus.list.df)%>%filter(File=="DonaldJTrump", type=="nomin")%>%select(sent.id, Economy:Legislation)
speech.df=as.matrix(speech.df)
speech.df[,-1]=replace(speech.df[,-1], speech.df[,-1]<1/15, 0.001)
speech.df[,-1]=f.smooth.topic(x=speech.df[,1], y=speech.df[,-1])
plot.stacked(speech.df[,1], speech.df[,topic.plot+1],
            xlab="Sentences", ylab="Topic share", main="Donald Trump, Nomination")
```

```{r, fig.width=3.3, fig.height=5}
# [1] "Economy"         "America"         "Defense"         "Belief"         
# [5] "Election"        "Patriotism"      "Unity"           "Government"     
# [9] "Reform"          "Temporal"        "WorkingFamilies" "Freedom"        
# [13] "Equality"        "Misc"            "Legislation"       


par(mfrow=c(5, 1), mar=c(1,1,2,0), bty="n", xaxt="n", yaxt="n")


topic.plot=c(1, 13, 14, 15, 8, 9, 12)
print(topics.hash[topic.plot])

speech.df=tbl_df(corpus.list.df)%>%filter(File=="GeorgeBush", type=="inaug", Term==1)%>%select(sent.id, Economy:Legislation)
speech.df=as.matrix(speech.df)
speech.df[,-1]=replace(speech.df[,-1], speech.df[,-1]<1/15, 0.001)
speech.df[,-1]=f.smooth.topic(x=speech.df[,1], y=speech.df[,-1])
plot.stacked(speech.df[,1], speech.df[,topic.plot+1],
             xlab="Sentences", ylab="Topic share", main="George Bush, inaugural Speeches")

speech.df=tbl_df(corpus.list.df)%>%filter(File=="WilliamJClinton", type=="inaug", Term==1)%>%select(sent.id, Economy:Legislation)
speech.df=as.matrix(speech.df)
speech.df[,-1]=replace(speech.df[,-1], speech.df[,-1]<1/15, 0.001)
speech.df[,-1]=f.smooth.topic(x=speech.df[,1], y=speech.df[,-1])
plot.stacked(speech.df[,1], speech.df[,topic.plot+1],
             xlab="Sentences", ylab="Topic share", main="William J Clinton, inaugural Speeches")

speech.df=tbl_df(corpus.list.df)%>%filter(File=="GeorgeWBush", type=="inaug", Term==1)%>%select(sent.id, Economy:Legislation)
speech.df=as.matrix(speech.df)
speech.df[,-1]=replace(speech.df[,-1], speech.df[,-1]<1/15, 0.001)
speech.df[,-1]=f.smooth.topic(x=speech.df[,1], y=speech.df[,-1])
plot.stacked(speech.df[,1], speech.df[,topic.plot+1],
             xlab="Sentences", ylab="Topic share", main="George W. Bush, inaugural Speeches")

speech.df=tbl_df(corpus.list.df)%>%filter(File=="BarackObama", type=="inaug", Term==1)%>%select(sent.id, Economy:Legislation)
speech.df=as.matrix(speech.df)
speech.df[,-1]=replace(speech.df[,-1], speech.df[,-1]<1/15, 0.001)
speech.df[,-1]=f.smooth.topic(x=speech.df[,1], y=speech.df[,-1])
plot.stacked(speech.df[,1], speech.df[,topic.plot+1],
             xlab="Sentences", ylab="Topic share", main="Barack Obama, inaugural Speeches")
```
```{r, fig.width=4, fig.height=5}
# [1] "Economy"         "America"         "Defense"         "Belief"         
# [5] "Election"        "Patriotism"      "Unity"           "Government"     
# [9] "Reform"          "Temporal"        "WorkingFamilies" "Freedom"        
# [13] "Equality"        "Misc"            "Legislation"       


par(mfrow=c(5, 1))

topic.plot=c(1, 13, 14, 15, 8, 9, 12)
print(topics.hash[topic.plot])

speech.df=tbl_df(corpus.list.df)%>%filter(File=="RonaldReagan", type=="farewell")%>%select(sent.id, Economy:Legislation)
speech.df=as.matrix(speech.df)
speech.df[,-1]=replace(speech.df[,-1], speech.df[,-1]<1/15, 0.001)
speech.df[,-1]=f.smooth.topic(x=speech.df[,1], y=speech.df[,-1])
plot.stacked(speech.df[,1], speech.df[,topic.plot+1],
             xlab="Sentences", ylab="Topic share", main="Ronald Reagan, Farewell Speeches")

speech.df=tbl_df(corpus.list.df)%>%filter(File=="GeorgeBush", type=="farewell")%>%select(sent.id, Economy:Legislation)
speech.df=as.matrix(speech.df)
speech.df[,-1]=replace(speech.df[,-1], speech.df[,-1]<1/15, 0.001)
speech.df[,-1]=f.smooth.topic(x=speech.df[,1], y=speech.df[,-1])
plot.stacked(speech.df[,1], speech.df[,topic.plot+1],
             xlab="Sentences", ylab="Topic share", main="George Bush, Farewell Speeches")

speech.df=tbl_df(corpus.list.df)%>%filter(File=="WilliamJClinton", type=="farewell")%>%select(sent.id, Economy:Legislation)
speech.df=as.matrix(speech.df)
speech.df[,-1]=replace(speech.df[,-1], speech.df[,-1]<1/15, 0.001)
speech.df[,-1]=f.smooth.topic(x=speech.df[,1], y=speech.df[,-1])
plot.stacked(speech.df[,1], speech.df[,topic.plot+1],
             xlab="Sentences", ylab="Topic share", main="William J. Clinton, Farewell Speeches")

speech.df=tbl_df(corpus.list.df)%>%filter(File=="GeorgeWBush", type=="farewell")%>%select(sent.id, Economy:Legislation)
speech.df=as.matrix(speech.df)
speech.df[,-1]=replace(speech.df[,-1], speech.df[,-1]<1/15, 0.001)
speech.df[,-1]=f.smooth.topic(x=speech.df[,1], y=speech.df[,-1])
plot.stacked(speech.df[,1], speech.df[,topic.plot+1],
             xlab="Sentences", ylab="Topic share", main="George W Bush, Farewell Speeches")


speech.df=tbl_df(corpus.list.df)%>%filter(File=="BarackObama", type=="farewell")%>%select(sent.id, Economy:Legislation)
speech.df=as.matrix(speech.df)
speech.df[,-1]=replace(speech.df[,-1], speech.df[,-1]<1/15, 0.001)
speech.df[,-1]=f.smooth.topic(x=speech.df[,1], y=speech.df[,-1])
plot.stacked(speech.df[,1], speech.df[,topic.plot+1],
             xlab="Sentences", ylab="Topic share", main="Barack Obama, Farewell Speeches")
```

```{r}
speech.df=tbl_df(corpus.list.df)%>%filter(type=="nomin", word.count<20)%>%select(sentences, Economy:Legislation)

as.character(speech.df$sentences[apply(as.data.frame(speech.df[,-1]), 2, which.max)])

names(speech.df)[-1]

```


```{r, fig.width=3, fig.height=3}
presid.summary=tbl_df(corpus.list.df)%>%
  filter(type=="inaug", File%in%sel.comparison)%>%
  select(File, Economy:Legislation)%>%
  group_by(File)%>%
  summarise_each(funs(mean))

presid.summary=as.data.frame(presid.summary)
rownames(presid.summary)=as.character((presid.summary[,1]))
km.res=kmeans(scale(presid.summary[,-1]), iter.max=200,
              5)
fviz_cluster(km.res, 
             stand=T, repel= TRUE,
             data = presid.summary[,-1],
             show.clust.cent=FALSE)
```

